High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-05-14 notebooks/linearly_separable.qmd
some working on cutting down pages

```{julia}
# Model parameters:
epochs = 100
batch_size = minimum([Int(round(n_obs/10)), 128])
n_hidden = 16
activation = Flux.swish
builder = MLJFlux.MLP(
    hidden=(n_hidden, n_hidden, n_hidden,), 
    σ=activation
)
n_ens = 5                                   # number of models in ensemble
_loss = Flux.Losses.logitcrossentropy       # loss function
_finaliser = x -> x                         # finaliser function
```

```{julia}
# JEM parameters:
𝒟x = Normal()
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=batch_size,
)
α = [1.0,1.0,1e-1]      # penalty strengths
```

@@

```{julia}
n_regen = 200
n_each = batch_size
for (mod_name, model) in model_dict
    K = length(counterfactual_data.y_levels)
    input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
    𝒟x = Uniform(extrema(counterfactual_data.X)...)
    𝒟y = Categorical(ones(K) ./ K)
    sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
    opt = ImproperSGLD()
    plts = []
    for target in levels(labels)
        target_idx = findall(levels(labels) .== target)[1]
        f(x) = logits(model, x)
        X̂ = sampler(f, opt; niter=n_regen, n_samples=n_each, y=target_idx)
        ex = extrema(hcat(MLJFlux.reformat(X),X̂), dims=2)
        xlims = ex[1]
        ylims = ex[2]
        x1 = range(1.0f0.*xlims...,length=10)
        x2 = range(1.0f0.*ylims...,length=10)
        p(x) = probs(model, x)
        plt = Plots.contour(
            x1, x2, (x, y) -> p([x, y][:,:])[target_idx], 
            fill=true, alpha=0.5, title="Target: $target", cbar=true,
            xlims=xlims,
            ylims=ylims,
        )
        Plots.scatter!(
            MLJFlux.reformat(X)[1,:], MLJFlux.reformat(X)[2,:], 
            color=Int.(labels.refs), group=Int.(labels.refs), alpha=0.5
        )
        Plots.scatter!(
            X̂[1,:], X̂[2,:], 
            color=repeat([target_idx], size(X̂,2)), 
            group=repeat([target_idx], size(X̂,2)), 
            shape=:star5, ms=10
        )
        savefig(plt, joinpath(output_images_path, "linearly_separable_generated_$(mod_name).png"))
        push!(plts, plt)
    end
    plt = Plots.plot(plts..., layout=(1, 2), size=(2*500, 400), plot_title=mod_name)
    display(plt)
end
```

@@

# Benchmark generators:
generator_dict = Dict(
    "Wachter" => WachterGenerator(λ=λ₁),
    "REVISE" => REVISEGenerator(λ=λ₁),
    "Schut" => GreedyGenerator(),
    "ECCCo" => ECCCoGenerator(λ=Λ),
    "ECCCo (no CP)" => ECCCoGenerator(λ=[λ₁, 0.0, λ₃]),
    "ECCCo (no EBM)" => ECCCoGenerator(λ=[λ₁, λ₂, 0.0]),
)
```

@@

M = model_dict["JEM"]
X = X isa Matrix ? X : Float32.(permutedims(matrix(X)))
factual_label =  levels(labels)[1]
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target =  levels(labels)[2]
factual = predict_label(M, counterfactual_data, x_factual)[1]

>>>>

```{julia}
# Model parameters:
epochs = 100
batch_size = minimum([Int(round(n_obs/10)), 128])
n_hidden = 16
activation = Flux.swish
builder = MLJFlux.MLP(
    hidden=(n_hidden, n_hidden, n_hidden,), 
    σ=activation
)
n_ens = 5                                   # number of models in ensemble
_loss = Flux.Losses.crossentropy       # loss function
_finaliser = Flux.softmax                   # finaliser function
```

```{julia}
# JEM parameters:
𝒟x = Normal()
𝒟y = Categorical(ones(output_dim) ./ output_dim)
sampler = ConditionalSampler(
    𝒟x, 𝒟y, 
    input_size=(input_dim,), 
    batch_size=50,
)
α = [1.0,1.0,1e-1]      # penalty strengths
```

@@

```{julia}
n_regen = 1000
n_each = batch_size
for (mod_name, model) in model_dict
    K = length(counterfactual_data.y_levels)
    input_size = size(selectdim(counterfactual_data.X, ndims(counterfactual_data.X), 1))
    𝒟x = Uniform(extrema(counterfactual_data.X)...)
    𝒟y = Categorical(ones(K) ./ K)
    sampler = ConditionalSampler(𝒟x, 𝒟y; input_size=input_size)
    opt = ImproperSGLD()
    plts = []
    for target in levels(labels)
        target_idx = findall(levels(labels) .== target)[1]
        f(x) = logits(model, x)
        X̂ = sampler(f, opt; niter=n_regen, n_samples=n_each, y=target_idx)
        ex = extrema(hcat(MLJFlux.reformat(X),X̂), dims=2)
        xlims = ex[1]
        ylims = ex[2]
        x1 = range(1.0f0.*xlims...,length=10)
        x2 = range(1.0f0.*ylims...,length=10)
        p(x) = probs(model, x)
        plt = Plots.contour(
            x1, x2, (x, y) -> p([x, y][:,:])[target_idx], 
            fill=true, alpha=0.5, title="Target: $target", cbar=true,
            xlims=xlims,
            ylims=ylims,
        )
        Plots.scatter!(
            MLJFlux.reformat(X)[1,:], MLJFlux.reformat(X)[2,:], 
            color=Int.(labels.refs), group=Int.(labels.refs), alpha=0.5
        )
        Plots.scatter!(
            X̂[1,:], X̂[2,:], 
            color=repeat([target_idx], size(X̂,2)), 
            group=repeat([target_idx], size(X̂,2)), 
            shape=:star5, ms=10
        )
        savefig(plt, joinpath(output_images_path, "linearly_separable_generated_$(mod_name).png"))
        push!(plts, plt)
    end
    plt = Plots.plot(plts..., layout=(1, 2), size=(2*500, 400), plot_title=mod_name)
    display(plt)
end
```

```{julia}
#| output: true
#| echo: false
#| label: fig-losses
#| fig-cap: "Illustration of the smooth size loss and the configurable classification loss."

X_plot = matrix(X)
temp = 0.1

for (mod_name, model) in model_dict
    p1 = Plots.contourf(model.model, model.fitresult, X_plot, labels; plot_set_loss=true, zoom=0, temp=temp)
    p2 = Plots.contourf(model.model, model.fitresult, X_plot, labels; plot_classification_loss=true, zoom=0, temp=temp, clim=nothing, loss_matrix=ones(2,2))
    display(Plots.plot(p1, p2, size=(800,320), plot_title=mod_name))
end
```

@@

opt = Flux.Optimise.Descent(0.05)

# Benchmark generators:
generator_dict = Dict(
    "Wachter" => WachterGenerator(λ=λ₁, opt=opt),
    "REVISE" => REVISEGenerator(λ=λ₁, opt=opt),
    "Schut" => GreedyGenerator(),
    "ECCCo" => ECCCoGenerator(λ=Λ, opt=opt),
    "ECCCo (no CP)" => ECCCoGenerator(λ=[λ₁, 0.0, λ₃], opt=opt),
    "ECCCo (no EBM)" => ECCCoGenerator(λ=[λ₁, λ₂, 0.0], opt=opt),
)
```

@@

M = model_dict["MLP"]
X = X isa Matrix ? X : Float32.(permutedims(matrix(X)))
factual_label =  levels(labels)[2]
x_factual = reshape(X[:,rand(findall(predict_label(M, counterfactual_data).==factual_label))],input_dim,1)
target =  levels(labels)[1]
factual = predict_label(M, counterfactual_data, x_factual)[1]

### 2023-05-14 notebooks/mnist.qmd
some working on cutting down pages

# Data:
n_obs = 10000
counterfactual_data = load_mnist(n_obs)
counterfactual_data.X = pre_process.(counterfactual_data.X)
counterfactual_data.generative_model = vae
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
x_factual = reshape(pre_process(x_factual, noise=0.0f0), input_dim, 1)
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
n_digits = Int(sqrt(input_dim))
output_dim = length(unique(labels))
```

@@

```{julia}
# Model parameters:
epochs = 100
batch_size = minimum([Int(round(n_obs/10)), 128])
n_hidden = 128
activation = Flux.swish
builder = MLJFlux.@builder Flux.Chain(
    Dense(n_in, n_hidden, activation),
    Dense(n_hidden, n_out),
)
n_ens = 5                                   # number of models in ensemble
_loss = Flux.Losses.logitcrossentropy       # loss function
_finaliser = x -> x                         # finaliser function
```

>>>>

# Data:
n_obs = 1000
counterfactual_data = load_mnist(n_obs)
counterfactual_data.X = pre_process.(counterfactual_data.X)
counterfactual_data.generative_model = vae
X, y = CounterfactualExplanations.DataPreprocessing.unpack_data(counterfactual_data)
X = table(permutedims(X))
x_factual = reshape(pre_process(x_factual, noise=0.0f0), input_dim, 1)
labels = counterfactual_data.output_encoder.labels
input_dim, n_obs = size(counterfactual_data.X)
n_digits = Int(sqrt(input_dim))
output_dim = length(unique(labels))
```

@@

```{julia}
# Model parameters:
epochs = 100
batch_size = minimum([Int(round(n_obs/10)), 128])
n_hidden = 128
activation = Flux.swish
builder = MLJFlux.@builder Flux.Chain(
    Dense(n_in, n_hidden, activation),
    Dense(n_hidden, n_out),
)
n_ens = 5                                   # number of models in ensemble
_loss = Flux.Losses.crossentropy       # loss function
_finaliser = Flux.softmax                        # finaliser function
```

### 2023-05-14 notebooks/poc.qmd
some working on cutting down pages

>>>>

```{julia}
#| output: true
#| echo: false
#| label: fig-losses
#| fig-cap: "Illustration of the smooth size loss and the configurable classification loss."

temp = 0.25
p1 = Plots.contourf(mach.model, mach.fitresult, permutedims(X), labels; plot_set_loss=true, zoom=0, temp=temp)
p2 = Plots.contourf(mach.model, mach.fitresult, permutedims(X), labels; plot_classification_loss=true, zoom=0, temp=temp, clim=nothing, loss_matrix=ones(2,2))
display(Plots.plot(p1, p2, size=(800,320)))
```
